---
title: "Statistical Inference Course Project Part 2"
author: "Chenchen"
date: "May 21, 2015"
output: pdf_document
---
## Overview

## 1. Load the ToothGrowth data and perform some basic exploratory data analyses 
```{r, echo=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
```

```{r}
# load data
data(ToothGrowth)
head(ToothGrowth)
```

```{r}
# change dose to factors
ToothGrowth$dose_fac = factor(ToothGrowth$dose, levels = c(0.5, 1, 2), labels = c("small", "medium", "large"))
head(ToothGrowth)
```
### Exploratory plot 1
```{r, echo=FALSE, fig.height=4}
qplot(supp, len, color = supp, geom = c("boxplot"), facets = .~ dose_fac, data = ToothGrowth)
```

### Exploratory plot 2
```{r, echo=FALSE}
qplot(dose_fac, len, color = dose_fac, geom = c("boxplot"), facets = supp ~., data = ToothGrowth)
```

## 2. Provide a basic summary of the data.
```{r}
summary(ToothGrowth)
```

## 3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. 
### 3.1 The distribution of treatment groups 
```{r, echo=FALSE}
# process data
ToothGrowth = ToothGrowth[, -3] # remove the original dose column
attach(ToothGrowth)
```
```{r}
tapply(len, list(dose_fac, supp),function(x) length(x)) 
```
```{r, echo=FALSE}
detach(ToothGrowth)
```

### 3.2 perform t-test of OJ vs VC
```{r}
# Subset data by dosage treatment
ToothGrowth_sm = filter(ToothGrowth, dose_fac == "small")
ToothGrowth_md = filter(ToothGrowth, dose_fac == "medium")
ToothGrowth_lg = filter(ToothGrowth, dose_fac == "large")
```

### 3.2.1 perform t-test of OJ vs VC in small dose group only
```{r}
t_test_sm = t.test(ToothGrowth_sm$len ~ ToothGrowth_sm$supp)
# confidence interval
t_test_sm$conf.int[1:2]
# p value
t_test_sm$p.value
```

### 3.2.2 perform t-test of OJ vs VC in medium dose group only
```{r}
t_test_md = t.test(ToothGrowth_md$len ~ ToothGrowth_md$supp)
# confidence interval
t_test_md$conf.int[1:2]
# p value
t_test_md$p.value
```

### 3.2.3 perform t-test of OJ vs VC in large dose group only
```{r}
t_test_lg = t.test(ToothGrowth_lg$len ~ ToothGrowth_lg$supp)
# confidence interval
t_test_lg$conf.int[1:2]
# p value
t_test_lg$p.value
```

### 3.3 perform t-test of small dosage (small) vs large dosage (large))
```{r}
# For convenience, we only compare small vs large, while drop medium dosages (medium)
ToothGrowth = filter(ToothGrowth, dose_fac != "medium")
ToothGrowth$dose_fac = factor(ToothGrowth$dose_fac)
# Subset data by delivery method
ToothGrowth_OJ = filter(ToothGrowth, supp == "OJ")
ToothGrowth_VC = filter(ToothGrowth, supp == "VC")
```

### 3.3.1 perform t-test of dose_sm vs dose_lg in OJ group only
```{r}
t_test_OJ = t.test(ToothGrowth_OJ$len ~ ToothGrowth_OJ$dose_fac)
# confidence interval
t_test_OJ$conf.int[1:2]
# p value
t_test_OJ$p.value
```

### 3.3.2 perform t-test of dose_sm vs dose_lg in VC group only
```{r}
t_test_VC = t.test(ToothGrowth_VC$len ~ ToothGrowth_VC$dose_fac)
# confidence interval
t_test_VC$conf.int[1:2]
# p value
t_test_VC$p.value
```

## 4. State your conclusions and the assumptions needed for your conclusions.

### 4.1 conclusions
#### 4.1.1 comparing delivery methods
- When treated with small and medium amount of Vitamin C, orange juice increases more tooth length than ascorbic acid.
- When treated with large amount of Vitamin C, there's no significant difference between the two delivery methods in influencing tooth length.

#### 4.1.2 comparing dose amount
- For both methods, large amount treatment will increase tooth length than small amount

### 4.2 Assumptions
- These tooth length are approximately normally distributed
- These samples are independent (unpaired)





